---
title: "Homework 2"
author: "Kelvin Delali Tamakloe"
date: 10/09/2021
output: github_document
---
```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)

```

# Problem 1

## Importing, reading and cleaning Mr.Trash Wheel sheet.

The code below will read the sheet of interest in the dataset, omit non-data entries
and rows that do not include dumpster-specific data, and round the number of sports
balls to the nearest integer.

```{r trashwheel}
trashwheel_df = 
  read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 1, range = "A2:N534") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, 0))
```

## Reading and cleaning the 2018 and 2019 precipitation data.

```{r precip}
precip2019_df = 
  read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 6, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019)

precip2018_df = 
  read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 7, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018)

precip_2018_2019_df =
  bind_rows(precip2019_df, precip2018_df) %>%
  mutate(month = month.name[month])
```

## Description of Data
The total number of observations in the Mr. Trash Wheel sheet is `r nrow(trashwheel_df)` observations. It also has `r ncol(trashwheel_df)` variables. Examples of the variables in the dataset are `r names(trashwheel_df %>% select(c(1, 3, 5)))`. In the data, there are dumpsters numbered from `r  min(pull(trashwheel_df, dumpster))` to `r  max(pull(trashwheel_df, dumpster))`. The median number of sports balls in a dumpster in 2019 is `r median(pull(filter(trashwheel_df, year == 2019), sports_balls))` balls.

The total number of observations in the precipitation data set is `r nrow(precip_2018_2019_df)` observations. It also has `r ncol(precip_2018_2019_df)` variables. Examples of the variables in the dataset are `r names(trashwheel_df %>% select(c(1, 2, 3)))`. 2018 had a total precipitation of `r sum(pull(filter(precip_2018_2019_df, year == 2018), total))` inches and 2019 had `r sum(pull(filter(precip_2018_2019_df, year == 2019), total))` inches.



