---
title: "Homework 1"
author: "Kelvin Delali Tamakloe"
date: 09/29/2021
output: github_document
---

Load the tidyverse library.

```{r setup}
library(tidyverse)
```

## Problem 1

The code chunk below produces a data frame that includes:
a random sample of size 10 from a standard Normal distribution, 
a logical vector indicating whether elements of the sample are greater than 0,
a character vector of length 10,
and a factor vector of length 10, with 3 different factor “levels”

```{r}
problem1_df = tibble(
  vec_random = rnorm(10),
  vec_logical = vec_random > 0,
  vec_charac = c("Kelvin", "is", "my", "name", "and", "data", "science", "is",
               "my", "game"),
  vec_factor = factor(c("down", "middle", "up", "up", "up", "middle", "down", "middle", "up", "down"))
)
```

#### To take the mean of each variable in problem1_df

```{r}
problem1_df
mean(pull(problem1_df, vec_random))
mean(pull(problem1_df, vec_logical))
mean(pull(problem1_df, vec_charac))
mean(pull(problem1_df, vec_factor))
```

Taking the mean of the numerical random sample vector (vec_random) and 
logical vector (vec_logical) works. Taking the mean of the 
character vector (vec_charac) and factor vector (vec_factor) does not work.
For the functions that did not work, a warning message stating "argument is 
not numeric or logical: returning NA" was returned.

#### To convert the variables in problem1_df to numeric

```{r numeric_convert, results = FALSE, warning = FALSE}
as.numeric(pull(problem1_df, vec_logical))
as.numeric(pull(problem1_df, vec_charac))
as.numeric(pull(problem1_df, vec_factor))
```

Converting the logical vector to a numeric vector with 0's and 1's works,
thus explaining our ability to obtain a mean for this. This happens 
as the logical vector is outputting TRUE or FALSE, which may be converted
converted to 0's and 1's. 

Converting the factor vector to a numeric vector with 1, 2, and 3 works.This
is because the software recognizes the factor vector has three levels and assigns 
each level a number.This reason does not account for a mean not being produced for 
the factor vector which perhaps may be due to a mean value perhaps not being a 
good statistic to describe all three levels. 

Converting the character vector to a numeric vector does not work. It also returns 
an NA value.This is as there is no underlying numeric integer to represent the 
character vector. This accounts for a mean not being generated for the character 
vector, as the character vector can not be converted to a numeric format from which
a mean may be calculated.

## Question 2

The following code loads the penguins dataset, includes a description 
of the penguins dataset, and provides a scatterplot of the data.

```{r}
data("penguins", package = "palmerpenguins")
flipper_mean = mean(pull(penguins,flipper_length_mm), na.rm = TRUE)
```

#### Description of Penguin Dataset

Below we describe the dataset including names, 
the size of the dataset, and the mean flipper length.


The data in this dataset includes `r colnames(penguins)`. The size of dataset is `r nrow(penguins) * ncol(penguins)`. The mean flipper length is `r flipper_mean`.

#### Scatterplot of Penguin Dataset

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
ggsave("scatter_plot.pdf", height = 4, width = 6)
```

