---
title: "Homework 2"
author: "Kelvin Delali Tamakloe"
date: 10/09/2021
output: github_document
---
```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)

```

# Problem 1

## Importing, reading and cleaning Mr.Trash Wheel sheet.

The code below will read the sheet of interest in the dataset, omit non-data entries
and rows that do not include dumpster-specific data, and round the number of sports
balls to the nearest integer.

```{r trashwheel}
trashwheel_df = 
  read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 1, range = "A2:N534") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, 0))
```

## Reading and cleaning the 2018 and 2019 precipitation data.

```{r precip}
precip2019_df = 
  read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 6, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019)

precip2018_df = 
  read_excel("Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 7, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018)

precip_2018_2019_df =
  bind_rows(precip2019_df, precip2018_df) %>%
  mutate(month = month.name[month])
```

## Description of Data
The total number of observations in the Mr. Trash Wheel sheet is `r nrow(trashwheel_df)` observations. It also has `r ncol(trashwheel_df)` variables. Examples of the variables in the dataset are `r names(trashwheel_df %>% select(c(1, 3, 5)))`. In the data, there are dumpsters numbered from `r  min(pull(trashwheel_df, dumpster))` to `r  max(pull(trashwheel_df, dumpster))`. The median number of sports balls in a dumpster in 2019 is `r median(pull(filter(trashwheel_df, year == 2019), sports_balls))` balls.

The total number of observations in the precipitation data set is `r nrow(precip_2018_2019_df)` observations. It also has `r ncol(precip_2018_2019_df)` variables. Examples of the variables in the dataset are `r names(trashwheel_df %>% select(c(1, 2, 3)))`. 2018 had a total precipitation of `r sum(pull(filter(precip_2018_2019_df, year == 2018), total))` inches and 2019 had `r sum(pull(filter(precip_2018_2019_df, year == 2019), total))` inches.



# Question 2

## Reading and cleaning the pols-month.csv file

```{r pols, message = FALSE}
polsmonth_df = 
  read_csv("pols-month.csv") %>% 
  janitor::clean_names() %>%
  
  separate(mon,into = c("year","month","day"),sep = "-") %>%
  pivot_longer(c(prez_gop,prez_dem), names_prefix = "prez_", names_to = "president") %>% 
  filter(value > 0) %>%
  
  select(-day, -value) %>% 
  
  mutate(year = as.integer(year), month = as.integer(month), month = month.name[month]) %>% 
  arrange(year, month)

```

## Reading and cleaning the snp.csv file

```{r snp, message = FALSE}
snp_df = 
  read_csv("snp.csv") %>% 
  janitor::clean_names() %>%
  
  separate(date,into = c("month","day","year"),sep = "/") %>%
  mutate(year = as.integer(year), year = ifelse(year > 21, 1900 + year, 2000 + year), month = as.integer(month), month = month.name[month]) %>% 
  
  select(-day) %>% 
  select(year, month, close) %>% 
  
  arrange(year, month)

```

## Reading and cleaning the unemployment.csv file

```{r unemploy, message = FALSE}
unemploy_df = 
  read_csv("unemployment.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer(c(jan:dec),
    names_to = "month", 
    values_to = "unemployment") %>%
  mutate(year = as.integer(year),
    month = recode(month, "jan" = "January", "feb" = "February", "mar" = "March", "apr" = "April", "may" = "May", "jun" = "June", "jul" = "July", "aug" = "August", "sept" = "September", "oct" = "October", "nov" = "November", "dec" = "December")) %>% 
  arrange(year, month)

```

## Merge and describe data
```{r merge, message = FALSE}
ans_merge_df =  left_join(polsmonth_df, snp_df, by = c("year", "month"))

ans_final_df = left_join(ans_merge_df, unemploy_df, by = c("year", "month"))

```

The polsmonth_df dataset contains `r nrow(polsmonth_df)` observations of `r ncol(polsmonth_df)` variables, that describe the national number of democratic and republican politicians at various points in time.

The snp_df dataset contains `r nrow(snp_df)` observations of `r ncol(snp_df)` variables regarding Standard & Poor’s stock market index (S&P), 

The merged dataset, ans_final_df, contains `r nrow(ans_final_df)` observations of `r ncol(ans_final_df)` variables. with a range of years from `r  min(pull(ans_final_df, year))` to `r  max(pull(ans_final_df, year))`. Some key variables include `r names(ans_final_df %>% select(c(1, 9, 10, 11)))`. 


